---
title: "Analyzing Respirometry Files from  Multiplexer Data"
author: "Julian Cassano"
date: "1/20/2022"
output: html_document
---
#Load in Packages
```{r}
library(dplyr)
library(stringr)
library(foreach)
library(segmented)
library(grDevices)
library(tidyr)
library(zoo)
library(tidyverse)
library(plotly)
library(hrbrthemes)
library(gridExtra)
library(ggpubr)
library(tidyverse)
library(readxl)
library(fs)
```

#Read in single bee Data files
```{r}
load.MR.file <- function(){
  file <- file.choose()
  filename <<- basename(file)
  Date <<- unlist(strsplit(filename, "[_]"))[1]
  Cohort <<-unlist(strsplit(filename, "[_]"))[2]
  Party <<- unlist(strsplit(filename, "[_]"))[3]
  Week <<- unlist(strsplit(filename, "[_]"))[4]
  Treatment <<- unlist(strsplit(filename, "[_]"))[5]
  Bee <<- unlist(strsplit(filename, "[_]"))[6]
  Bee <<- gsub(".csv", "", Bee)
  In.Data   <- read_csv(file)
  return(In.Data)
  #write.csv(restfull, "test.csv", row.names=F)
}
In.Data <- load.MR.file()

#Wrangle the data
In.Data <-  In.Data%>%
  dplyr::rename("Carbon_Dioxide" = "CO2",
         "Oxygen" = "O2")%>%
  dplyr::select(-c("Temp__Aux2_", "FoxBoxTemp", "FlowSet"))%>% #select variables of interest
  mutate(markers = ifelse(Mux__Aux1_ == "8", "baseline", "MR"),
         Total_Seconds = seq.int(nrow(In.Data))) #Add an indexing column that numbers each row 1: number of rows
str(In.Data)
View(In.Data)
```

#Determining beginning of baseline and chamber MR measurements
```{r}
#Saving baseline and MR measurements to vector objects
bsln <- which(In.Data$markers == "baseline") #Baseline at beginning
chbr <- which(In.Data$markers == "MR") #Indicates the switch between baseline and

#Determining the start of each vector
bsln1 <- bsln[10]
chbr1 <- chbr[1]
```

#Visualize CO2 and O2
```{r}
p1 <- In.Data%>%
  ggplot( aes(x=Total_Seconds, y=Carbon_Dioxide)) +
    geom_line() +
    ylab("CO2") +
    theme_ipsum()
p2 <- ggplotly(p1)
p2

p3 <- In.Data%>%
  ggplot( aes(x=Total_Seconds, y=Oxygen)) +
    geom_line()+
    ylab("O2") +
    theme_ipsum()
p4 <- ggplotly(p3)
p4
```

#Mean Baseline determination for correction
```{r}
#Subset dataframe with just baseline values
Baseline.df <- dplyr::filter(In.Data, Total_Seconds < chbr1, Total_Seconds > bsln1 )
#Calculate the average baseline value
mean_base_CO2 <- mean(Baseline.df$Carbon_Dioxide)
mean_base_O2 <- mean(Baseline.df$Oxygen)

#Correct Carbon_Dioxide values for mean baseline
In.Data<- In.Data%>%
  mutate(Carbon_Dioxide_pct = Carbon_Dioxide + abs(mean_base_CO2), #subtract the mean CO2 baseline from the CO2 readings from the foxbox
         Oxygen_pct = mean_base_O2 - Oxygen) #gives the oxygen CONSUMED
# units still %CO2
# units still %O2
```

#Visualize only timepoints where the bee is in the chamber
```{r}
#CO2
p1 <- In.Data%>%
  ggplot( aes(x=Total_Seconds, y=Carbon_Dioxide_pct)) +
    geom_line()+
    ylab("Percent CO2 Consumed") +
    theme_ipsum()
p2 <- ggplotly(p1)
p2

#O2
p3 <- In.Data%>%
  ggplot( aes(x=Total_Seconds, y=Oxygen_pct)) +
    geom_line()+
    ylab("Percent O2 Consumed") +
    theme_ipsum()
p4 <- ggplotly(p3)
p4

ggarrange(p1 , p3,  ncol = 1, nrow = 2)
```

#Add VCO2 to data frame and calculate output values
```{r}
# Get VCO2 by multiplying by the flow rate (in ml/min)
In.Data<-In.Data%>%
  dplyr::mutate(VCO2 = In.Data$Carbon_Dioxide_pct / 100 * In.Data$Flow_Rate * 60)
# units are NOW :: ml CO2 / hr and are fully corrected 
##this is essentially equation 10.3 from Lighton 2008 book
In.Data<-In.Data%>%
  dplyr::mutate(VO2 = In.Data$Oxygen_pct / 100 * In.Data$Flow_Rate * 60) # mL O2 / hr 

In.Data.MR<-In.Data%>%
  filter(Total_Seconds > chbr1 & Total_Seconds < chbr1+(60*25)) # make "usetime" only MR measurements
```

#Visualize just MR data
```{r}
#GGplot for CO2 with only MR measurements (no baseline)
p5 <- In.Data.MR%>%
  ggplot( aes(x=Total_Seconds, y=VCO2)) +
    geom_line()+
    ylab("VCO2") +
    scale_alpha_manual(values= c(1,.1))+
    theme_ipsum()
p6 <- ggplotly(p5)
p6
```

#Summary Statistics
```{r}
#This creates what is essentially a row in the final data set
sum.stats<-In.Data.MR %>%
  dplyr::summarise(
    Filename = filename,
    Date = Date,
    Cohort = Cohort,
    Party = Party,
    Week = Week,
    Treatment = Treatment,
    Bee = Bee,
    count = n(),
    meanVCO2 = mean(VCO2, na.rm = TRUE),
    sdVCO2 = sd(VCO2, na.rm = TRUE),
    seVCO2 = sd(VCO2) / sqrt((length(VCO2))))
view(sum.stats)
```

#Write a master csv file with each row comntaining values from a single data file
```{r}
#Create a CSV to populate with rows created for each file above. This originates by making the csv in excel
MR_master <- read.csv("MR_Exp1_Master.csv", header = T, sep=",") # Change this to the desired file name for your application
MR_master$Filename <- as.character(MR_master$Filename)
str(MR_master)

#Add newest processed row to previous data, overwrite the csv (then move on to the next one)
MR_master <- rbind(MR_master, sum.stats)
write.csv(MR_master, "MR_Exp1_Master.csv", row.names=F)

# check then clear the workspace for the next processing
#ls()
rm(list = ls())
```



